Query Optimization for Dynamic Imputation

### 2017-05-15 paper/experiments.tex
Add numbers to discussion of other imputation algs

\review{We evaluate \ProjectName{} optimizing for quality ($\alpha=0$), optimizing for performance
($\alpha=1$), and targeting a balance of each ($\alpha = 0.5$). As a baseline, we fully impute the tables
used by each query and then run the query on the imputed tables, simulating the process that
an analyst would follow to apply existing imputation techniques.}

@@

\review{
We also experiment with mean-value imputation and hot
deck imputation (\Cref{sec:imputation}). In our implementation of mean-value imputation,
to facilitate non-blocking operation within the iterator
model, we estimate the column mean from base table histograms. Therefore,
imputations are identical no matter where the operator is placed in the query
plan. Moreover, runtimes are essentially identical regardless of the value of
$\alpha$ because the time cost of the imputation is relatively low independent of
how many tuples need to be imputed. In hot deck imputation, runtimes are also
similar across $\alpha$, because the cost of buffering
tuples and sampling at random is negligible compared to the rest of query
execution. \todo{be quantitative: In terms of accuracy, relative errors using hot deck are close to
zero; since the baseline, in this case, is hot deck imputation on the base
table, only skew in the distributions arising from filters and joins would lead
to larger errors. In some cases, skew does arise, and errors are on the same
magnitude as in the decision trees case. Indeed, \ProjectName{} yields larger
benefit when using relatively sophisticated, higher quality imputation
algorithms as opposed to simpler strategies that would even be quick to
implement on the base table.
}
}

@@

\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced by the planner for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm
and encapsulate the possible trade-offs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to guide their workflow.
}

>>>>

We evaluate \ProjectName{} optimizing for quality ($\alpha=0$), optimizing for performance
($\alpha=1$), and targeting a balance of each ($\alpha = 0.5$). As a baseline, we fully impute the tables
used by each query and then run the query on the imputed tables, simulating the process that
an analyst would follow to apply existing imputation techniques.

@@

\review{
We also experiment with mean-value imputation and hot deck imputation
(\Cref{sec:imputation}). In our implementation of mean-value imputation, to facilitate
non-blocking operation within the iterator model, we estimate the column mean from base
table histograms. Therefore, imputed values are identical no matter where the operator is
placed in the query plan (relative errors are $0$). Runtimes are in the range of
\runtimemeanlow{} to \runtimemeanhigh{} for queries using \ProjectName{}, as compared to
\runtimemeanlowbase{} to \runtimemeanhighbase{} when imputing on the base table.  Here, the
time cost of the imputation is relatively low no matter how many tuples need to be imputed.

In hot deck imputation, runtimes are also similar across $\alpha$, ranging from
\runtimehotdecklow{} to \runtimehotdeckhigh{} using dynamic imputation, as compared to
\runtimehotdecklowbase{} to \runtimehotdeckhighbase{} when imputing on the base table. In
this case, the cost of buffering tuples and sampling at random is negligible compared to the
rest of query execution. Relative errors using hot deck are close to zero, ranging from
\smapehotdecklow{}--\smapehotdeckhigh{}\%. (The exception is \Cref{q8} with $\alpha=1$, which has error
of \smapehotdeckhighoutlier{}\% and exhibits significant skew.) Since the baseline, in these
cases, is hot deck imputation on the base table, it is skew in the distributions arising from
filters and joins that often leads to larger errors. Indeed, \ProjectName{} yields
larger benefit when using relatively sophisticated, higher quality imputation algorithms as
opposed to simpler strategies that would even be quick to implement on the base table.
}

@@

\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced by the planner for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm
and encapsulate the possible trade-offs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to help in guiding their workflow.
}

### 2017-05-15 paper/queries.tex
added missing predicate to Q8 for queries table
Pass through entire paper
Add numbers to discussion of other imputation algs

\newcounter{queryno}
{
  \lstset{breaklines=true, basicstyle=\scriptsize\ttfamily}
\begin{tabular}{cl}
  \toprule
  \# & \multicolumn{1}{c}{Queries on CDC data} \\
  \midrule
  1 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q1} \\[-1.5ex]
  2 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(creatine) FROM demo, exams, labs WHERE demo.id = exams.id AND exams.id = labs.id AND income >= 13 AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
      \end{minipage}
      \refstepcounter{queryno}\label[query]{q2} \\[-1.5ex]
  3 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q3}\\[-1.5ex]
  4 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q4}\\[-1.5ex]
  5 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
  \midrule
  \# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
  \midrule
  6 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q6} \\[-1.5ex]
  7 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q7} \\[-1.5ex]
  8 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc WHERE studentdebtowe > 0 GROUP BY schooldegree;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q8}\\[-1.5ex]
  9 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp 
WHERE fcc.countrycitizen = gdp.country AND age >= 18 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q9}\\[-1.5ex]
  \bottomrule
\end{tabular}
}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\newcounter{queryno}
{
  \lstset{breaklines=true, basicstyle=\scriptsize\ttfamily}
\begin{tabular}{cl}
  \toprule
  \# & \multicolumn{1}{c}{Queries on CDC data} \\
  \midrule
  1 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q1} \\[-1.5ex]
  2 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(creatine) FROM demo, exams, labs
WHERE demo.id = exams.id AND exams.id = labs.id
    AND income >= 13AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
      \end{minipage}
      \refstepcounter{queryno}\label[query]{q2} \\[-1.5ex]
  3 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q3}\\[-1.5ex]
  4 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q4}\\[-1.5ex]
  5 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
  \midrule
  \# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
  \midrule
  6 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q6} \\[-1.5ex]
  7 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q7} \\[-1.5ex]
  8 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc
WHERE studentdebtowe > 0 and schooldegree >= 0 GROUP BY schooldegree;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q8}\\[-1.5ex]
  9 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp 
WHERE fcc.countrycitizen = gdp.country AND age >= 18 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q9}\\[-1.5ex]
  \bottomrule
\end{tabular}
}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
